Statistics Project

### 2022-12-12 StatsProject_draft_Caro.Rmd
Custom Model Trials

## Read in data
```{r}
train <- read.csv("train.csv")
test <- read.csv("test.csv")

@@

head(train_a1)
head(test_a1)
```
## Summary Statistics
```{r}
summary(train_a1)
summary(test_a1)
```
## EDA 
```{r}
train_a1 %>% ggplot(aes(x = (GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

@@

# Linear Model
```{r}
fit <- lm(log(SalePrice) ~ log(GrLivArea) + Neighborhood + GrLivArea*Neighborhood, data = train_a1)
summary(fit)
confint(fit)

plot(fit)
```
# ANOVA 
```{r}
res.aov <- aov(log(SalePrice)~log(GrLivArea)+Neighborhood+Neighborhood*GrLivArea, data = train_a1)
summary(res.aov)
```
```{r}
library(car)
# QQPlot
qqPlot(log(train_a1$SalePrice))

# histogram of residuals 
h <- hist(fit$residuals, breaks = 10, density = 10,col = "lightgray",main = "Residuals") 
xfit <- seq(min(fit$residuals), max(fit$residuals), length = 40) 
yfit <- dnorm(xfit, mean = mean(fit$residuals), sd = sd(fit$residuals)) 
yfit <- yfit * diff(h$mids[1:2]) * length(fit$residuals) 
```

>>>>

## Read in data
```{r}
train <- read.csv("train.csv", header=TRUE, stringsAsFactors=TRUE)
test <- read.csv("test.csv", header=TRUE, stringsAsFactors=TRUE)

@@

head(train_a1)
head(test_a1)
```
## Summary Statistics
```{r}
summary(train_a1)
summary(test_a1)

str(train_a1)
```
## EDA

```{r}
train_a1 %>% ggplot(aes(x = (GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

@@

## Linear Model 
```{r}
fit1 <- lm(log(SalePrice) ~ log(GrLivArea) + Neighborhood + GrLivArea*Neighborhood, data = train_a1)
summary(fit1)
confint(fit1)

plot(fit1)

hist(fit1$residuals, col = "blue", main = "Histogram of Residuals")
plot(fit1$fitted.values,fit1$residuals, main = "Plot of Residuals v. Fitted Values")
sqrt(mean(fit1$residuals^2)) #RMSE
```
```{r}

```

## ANOVA 
```{r}
res.aov <- aov(log(SalePrice)~log(GrLivArea)+Neighborhood+Neighborhood*GrLivArea, data = train_a1)
summary(res.aov)
```
```{r}
library(car)
# QQPlot
qqPlot(log(train_a1$SalePrice))

# histogram of residuals 
h <- hist(fit1$residuals, breaks = 10, density = 10,col = "lightgray",main = "Residuals") 
xfit <- seq(min(fit1$residuals), max(fit1$residuals), length = 40) 
yfit <- dnorm(xfit, mean = mean(fit1$residuals), sd = sd(fit1$residuals)) 
yfit <- yfit * diff(h$mids[1:2]) * length(fit1$residuals) 
```

##remove outliers
```{r}

```

# ANALYSIS 2: MODELS

## EDA
```{r}
str(train)
summary(train)
```
##
### Missing data & Duplicates
- NA means something in following: Alley, BsmtQual, BsmtCond, BsmtExposure, BsmtFinType1,BsmtFinType2, FireplaceQu, GarageType, GarageFinish, GarageQual, GarageCond, PoolQC, Fence, MiscFeature
- Electrical has NAs but seems important? 
```{r}
colSums(is.na(train)) != 0

# remove columns with NAs that have no meaning (from data description)
train_clean = subset(train, select= -c(LotFrontage,MasVnrType, MasVnrArea))

```

```{r}
train[duplicated(train),]

```
## Look at distribution of SalePrice &  LotArea

## Replace N/A 
```{r}

train_clean[is.na(train_clean)] = "None"
summary(train_clean )
```

##  Feature Selection 
```{r}
library(caret)
# Feature Selection
# We need to find features which are important for the model.

# Setting Seed to ensure results are repeatable.
set.seed(7)

# prepare training scheme
control <- trainControl(method="repeatedcv", number=10, repeats=3)
# train the model 
model <- train(SalePrice~., data=train, method="lvq", preProcess="scale", trControl=control)
# estimate variable importance
importance <- varImp(model, scale=FALSE)
# summarize importance
print(importance)
# plot importance
plot(importance)
```
## Create Model 
```{r}
fit2 = lm(SalePrice~LotArea , data= train)
summary(fit2)
```

## Forward Selection

```{r}
library(olsrr)
ols_step_forward_p(fit2, penter = 0.05, details = TRUE)

```

## Backwards Elimination 
```{r}
ols_step_backward_p(fit, prem = 0.05, details = TRUE)

```

## Stepwise Selection 
```{r}
ols_step_both_p(fit, prem = 0.05, pent = 0.05, details = FALSE)

```

## Custom Model

### 2022-12-12 rough_analysis.Rmd
Custom Model Trials

For the first analysis, we are looking at houses are are only within the locale of [NAmes, Edwards, BrkSide neighborhoods] and at how Sale Price can possibly change
with respect to the Size of the living area and the Neighborhood in which the property is located, which can be visualized with the equation:

@@

a1_set %>% ggplot(aes(x = SalePrice/10000, fill = Neighborhood)) + geom_histogram(bins =20)
```

@@

#RMSE: 
ModelRMSE(fit$residuals)
```

@@

```{r}
#bad approach, maybe, I can't get it to work 
#train.control = trainControl(method = "cv", number = 10)
#step.model = train(SalePrice ~.-GarageArea, data = housing %>% select_if(is.numeric) %>% drop_na(), method = "leapBackward", trControl = train.control)
#step.model$modelInfo
```

@@

```{r}
#install.packages("olsrr")

@@

summary(transForward$model)
ModelRMSE(transForward$model$residuals) #see defined functions at top of document
CVPress(transForward$model)

@@

mystery4thsetWIP

@@

>>>>

## Restatement of the Problem 
Estimate how SalesPrice of the house is related to the square footage of the living area of the house & if SalesPrice (and its relationship to square footage) depends on which neighborhood the house is located in. We are only interested at looking at houses in the NAmes, Edwards & BrkSide neighborhoods. 
This relationship can be visualized with the equation:

@@

a1_set %>% ggplot(aes(x = SalePrice/10000, fill = Neighborhood)) + geom_histogram(bins =20)

a1_set %>% ggplot(aes(x = GrLivArea, fill = Neighborhood)) + geom_histogram(bins =20)

```
The histograms show both GrLivArea and Sales Price having a right skew distribution. We will examine different log transformations to see if we can obtain a more normal distribution suitable for our model.

## Log Transforms 
```{r}
# histograms with log transform on each individual variables
train_a1 %>% ggplot(aes(x = log(SalePrice/10000), fill = Neighborhood)) + geom_histogram(bins =20)

train_a1 %>% ggplot(aes(x = log(GrLivArea), fill = Neighborhood)) + geom_histogram(bins =20)

train_a1 %>% ggplot(aes(x = log(GrLivArea), y = SalePrice/1000, color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

train_a1 %>% ggplot(aes(x = (GrLivArea), y = log(SalePrice/1000), color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")

train_a1 %>% ggplot(aes(x = log(GrLivArea), y = log(SalePrice/1000), color = Neighborhood)) + geom_point() + stat_smooth(method = "lm")
```
Using a log transform on the GrLivArea and SalePrice variables, we achieve a more normal distribution for each variable which will be more conducive to use in our model.

@@

#RMSE: 
ModelRMSE(fit$residuals)
```
## Linear Models

From a visual inspection of the Cook's D and residual plot, we notice there are 3 outliers: 
  - ID 54
  - ID 61, 
  - ID 114
```{r}
fit1 <- lm(SalePrice ~ GrLivArea + Neighborhood + GrLivArea*Neighborhood, data = a1_set)
summary(fit1)
confint(fit1)

plot(fit1)

hist(fit1$residuals, col = "blue", main = "Histogram of Residuals")
sqrt(mean(fit1$residuals^2)) #RMSE
```

### Remove Outliers 
```{r}
head(a1_set)
a1_set2 = housing
a1_set2 = a1_set2 [(!(a1_set2 $Id == "54") & !(a1_set2 $Id == "61") & !(a1_set2 $Id == "114") ),]
head(a1_set2)

a1_set2 = a1_set2 %>% select(SalePrice, Neighborhood, GrLivArea) %>% filter( Neighborhood == c("NAmes", "Edwards", "BrkSide"))
head(a1_set2)
```

### Fit model without outliers 
```{r}
fit2 <- lm(log(SalePrice) ~ log(GrLivArea) + Neighborhood + GrLivArea*Neighborhood, data = a1_set2)
summary(fit2)
confint(fit2)

plot(fit2)

hist(fit2$residuals, col = "blue", main = "Histogram of Residuals")
sqrt(mean(fit2$residuals^2)) #RMSE
```

@@

```{r}
#bad approach, maybe, I can't get it to work 
train.control = trainControl(method = "cv", number = 10)
step.model = train(SalePrice ~.-GarageArea, data = housing %>% select_if(is.numeric) %>% drop_na(), method = "leapBackward", trControl = train.control)
pred(step.model, housing)
#step.model$modelInfo
```

@@

```{r}
install.packages("olsrr")

@@

transForward
summary(transForward$model)
ModelRMSE(transForward$model$residuals) #see defined functions at top of document
CVPress(transForward$model)

@@

forwardt1
backt1
stepwiset1
#mystery4thsetWIP

@@

```
## Custom Model 
```{r}
# Change categorical values to numeric
housing2 = housing
housing2$Neighborhood <- as.numeric(factor(housing2$Neighborhood,
                                                        levels = c("Blmngtn", "Blueste", "BrDale"),
                                                        labels = c(2,1,0), ordered = TRUE))
housing2$LotConfig <- as.numeric(factor(housing2$LotConfig,
                                                        levels = c("Inside", "Corner", "CulDSac", "FR2", "FR3"),
                                                        labels = c(4,3,2,1,0), ordered = TRUE))
housing2$ExterQual <- as.numeric(factor(housing2$ExterQual,
                                                        levels = c("Ex", "Gd", "TA", "Fa", "Po"),
                                                        labels = c(4, 3, 2,1,0), ordered = TRUE))
housing2$BsmtQual <- as.numeric(factor(housing2$BsmtQual,
                                                        levels = c("Ex", "Gd", "TA", "Fa", "Po"),
                                                        labels = c(4, 3, 2,1,0), ordered = TRUE))

```
```{r}
# Only look at Normal Sales 
normal_housing = housing 
normal_set = normal_housing %>% select(SaleCondition) %>% filter( SaleCondition == c("Normal"))
normal_housing = normal_housing %>% select_if(is.numeric)
```

```{r}
t2 = subset(transformed_numeric, select = -c(LotFrontage, MasVnrArea, GarageYrBlt, OverallQual, TotRmsAbvGrd, GarageCars, TotalBsmtSF, X1stFlrSF, X2ndFlrSF, FullBath )) 
#removal of largely correlated values to reduce multicollinearity (VIF /approx 1) and actually allow back-elimination to run,
#it wasn't running without a ton of errors without this.

corrplot(cor(t2)) #visualize correlation to identify which plots are correlated. After elimiation, it should be fairly sparsely correlated.

t2fit = lm(SalePrice ~. , data = t2)

backt2 = ols_step_backward_p(t2fit, prem = 0.05, details = FALSE)
forwardt2 = ols_step_forward_p(t2fit, penter = 0.05, details = FALSE)               #the p-values for these should be fiddled with, 
stepwiset2 = ols_step_both_p(t2fit, pent = 0.05, prem =0.05, details = FALSE)

forwardt2
backt2
stepwiset2
#mystery4thsetWIP

#---------------------------------Summary Stats of the produced RL sets
summary(backt2$model)
vif(backt2$model)

ModelRMSE(backt2$model$residuals)
CVPress(backt2$model)
#---------------------------------
summary(forwardt2$model)
vif(forwardt2$model)

ModelRMSE(forwardt2$model$residuals)
CVPress(forwardt2$model)
#---------------------------------
summary(stepwiset2$model)
vif(stepwiset2$model)

ModelRMSE(stepwiset2$model$residuals)
CVPress(stepwiset2$model)
